Statistical Machine Learning Models on A Bank Marketing dataset

### 2025-12-08 Slides.Rmd
eda slides

## The Bank Marketing Dataset (Slide 2)
* **Discuss:** The dataset structure: $\approx 45,000$ client observations from a Portuguese banking institution's telemarketing campaigns.
* **Discuss:** Brief overview of the four main predictor categories: Client Demographics, Financial Indicators, Campaign Features, and Past Outcomes. Mention the paper had more features, fro privacy reasons the public data set has less
* **Discuss:** The Target Variable ($y$) is binary: Did the client subscribe to a term deposit (Yes/No)?
* **Discuss:** Emphasize the severity of the class imbalance (e.g., typically $\approx 11\%$ 'Yes' subscriptions). This may yield higher false negatives.
* **Image/Plot Idea:** A small table summarizing the list of the 4 key variable groups (e.g., Demographics, Financial, Campaign, Past). 
* **Image/Plot Idea:** A **Bar Plot** showing the distribution of the target variable ($y$), visually highlighting the imbalance.

## Problem Statement & Methodology Overview (Slide 3)
* **Discuss:** The project aims for a **Comparative Study**—comparing models based on two criteria: **Predictive Power** and **Interpretability**.
* **Discuss:** The project goal is **Binary Classification**: Building statistical models that are both **predictive** and **interpretable** for management.
* **Discuss:** Established an **80/20 Train/Test Split** and used **5-Fold Cross-Validation (CV)** on the training set for tuning and estimating out-of-sample error.
* **Image/Plot Idea:** An icon representing **Predictive Power** next to an icon representing **Interpretability** (The two core goals).

>>>>

## The Bank Marketing Dataset

* **45,000 client observations** from Portuguese bank telemarketing  
* **Target variable ($y$):** Did client subscribe to term deposit? (Yes/No)  
* **Class imbalance:** Only ≈11% "Yes" subscriptions  
* **Four main predictor groups:** demographics, financial indicators, campaign features, past outcomes

## Term Deposit Class Imbalance

* **Most clients say NO; only ≈11% say YES**

```{r slide2-plot, echo=FALSE, fig.align="center", out.width="90%"}
knitr::include_graphics("imbalance_plot.png")
```

## EDA Finding 1: Contact Method Matters

```{r eda1-plot, echo=FALSE, fig.align="center", out.width="90%"}
knitr::include_graphics("contact_plot.png")

## EDA Finding 2: Previous Success Predicts Future Success

```{r eda2-plot, echo=FALSE, fig.align="center", out.width="90%"}
knitr::include_graphics("poutcome_plot.png")
```

## EDA Finding 3: Too Many Calls Hurts Results

```{r eda3-plot, echo=FALSE, fig.align="center", out.width="90%"}
knitr::include_graphics("campaign_plot.png")
```

## Problem Statement & Methodology Overview

* **Goal:** Compare models on two criteria:  
  * **Predictive Power** – ability to correctly find YES cases  
  * **Interpretability** – model is easy to understand

* **Model evaluation:**  
  * **80/20 Train/Test Split**  
  * **5-Fold Cross-Validation**
